---
title: "Exploring wildfire and wells data in Oklahoma"
subtitle: "August 2022"
output: html_notebook
---

## Memo

Compares two methods Leo used to determine which wells were within wildfire boundaries in Oklahoma. The '0km' method was restricted to wells within an individual state and may have under-counted fires. The '0km_plus' method accounted for wells in neighboring states, for wildfires that crossed state boundaries (and should have more wells than the 0km method) 

Here, I simply verify that there are more wells in the 'plus' wildfire exposure data.

```{r data_input, include=FALSE}
# imports data
fire_ok_all_dates_0km <- 
  readRDS("../../data/interim/fire_ok_all_dates_0km.rds")
fire_ok_all_dates_0km_plus <- 
  readRDS("../../data/interim/fire_ok_all_dates_0km_plus.rds")
```


```{r echo=FALSE}
# combines datasets
fires_ok <- fire_ok_all_dates_0km_plus %>% 
  select("Event_ID", "wells_ok_all_dates_0km") %>% 
  rename("wells_ok_all_dates_0km_plus" = "wells_ok_all_dates_0km") %>% 
  left_join(fire_ok_all_dates_0km, by = "Event_ID")
fires_ok %>% 
  select("Event_ID", "wells_ok_all_dates_0km", "wells_ok_all_dates_0km_plus") %>%
  mutate(difference = wells_ok_all_dates_0km_plus - wells_ok_all_dates_0km) %>% 
  arrange(desc(difference))
```

It looks like this worked. Two fires had more wells in the 'plus' method and no fires had fewer wells.I'll confirm that the fire with Event_ID == "OK3677410017520170307" originated in OK and crossed state boundaries.

This was the STARBUCK fire in 2017. According to news reports, this fire burned in the Oklahoma panhandle and southern Kansas, and was highly destructive (though I didn't see a mention of oil wells).

We can do further checks by mapping and visually inspecting the data.