---
title: "Exploring duplicate wildfires data"
subtitle: "December 2022"
output: html_notebook
---

## Memo

In current draft Figure 1 there appears to be duplicate wildfires in Kansas and Oklahoma. In this memo I explore those data to confirm whether there are in fact duplicates and, if so, will seek to resolve the issue.


```{r setup, include=FALSE}
source("../../code/0-setup/01-setup.R")
```

```{r data_input, include=FALSE}
wells_all     <- readRDS("../../data/processed/wells_all.rds")
wildfires_all <- readRDS("../../data/processed/wildfires_all.rds")
```


**Plot. Wildfires in Kansas and Oklahoma**

```{r echo=FALSE}
wildfires_ks_ok <- wildfires_all %>%
  filter(state %in% c("KS", "OK")) %>% 
  st_as_sf()
plot(st_geometry(wildfires_ks_ok))
```

Question - are these potentially duplicate wells potentially of relevance in the context of this study, by which I mean, are they near the wells in Kansas?

```{r echo=FALSE}
wildfires_ks <- wildfires_all %>% filter(state == "KS") %>% st_as_sf()
wells_ks <- wells_all %>% filter(state == "KS")

plot(st_geometry(wells_ks), pch = 4)
plot(st_geometry(wildfires_ks), add = T, col = "red")
```
There are a lotof wells in KS it seems (n = 55,626) and many of them are in the eastern part of the state where we're seeing this potential duplication issue. So it is salient!

Next step is to see if there is a way to identify duplicate geometries using dplyr.

```{r echo=FALSE}
wildfires_ks_ok2 <- wildfires_ks_ok %>% 
  as.data.frame() %>% 
  distinct(geometry)
```

This did result in two fewer wildfires (3,261 vs. 3,259 after distinct). Let's try this using sf methods.

```{r echo=FALSE}
wildfires_ks_ok3 <- wildfires_ks_ok %>% 
  mutate(eq     = st_equals(wildfires_ks_ok$geometry)) %>% 
  mutate(eq_dup = duplicated(eq))
```

Same answer as above; 2 duplicate fires. Let's take a look at them.

```{r echo=FALSE}
wildfires_ks_ok3 %>% filter(eq_dup == T) %>% View
```


