---
title: "Results: Wildfires in areas with projected wildfire risk"
output:
  html_notebook:
    theme: flatly
    highlight: haddock 
    toc: yes
    toc_depth: 2
    toc_float: yes
    number_sections: true
---

```{r setup, include=FALSE}
# loads necessary packages and defines global variables
source("../../code/0-setup/01-setup.R")
library("viridis")
```

```{r data_input, include=FALSE}
wells_kbdi <- readRDS("../../data/processed/wells_kbdi.rds")
study_region_sf <- readRDS("../../data/interim/study_region_sf.rds")
kbdi_max_2017_sf <- readRDS("../../data/processed/kbdi_max_2017.rds") %>% 
  as_tibble() %>%
  st_as_sf(coords = c("x", "y"), crs = crs_nad83) %>% 
  st_intersection(study_region_sf)
kbdi_max_2050_sf <- readRDS("../../data/processed/kbdi_max_2050.rds") %>% 
  as_tibble() %>%
  st_as_sf(coords = c("x", "y"), crs = crs_nad83) %>% 
  st_intersection(study_region_sf)
kbdi_max_2090_sf <- readRDS("../../data/processed/kbdi_max_2090.rds") %>% 
  as_tibble() %>%
  st_as_sf(coords = c("x", "y"), crs = crs_nad83) %>%
  st_intersection(study_region_sf)
us_states_west <- st_read("../../data/raw/us_census/tl_2018_us_state.shp") %>% 
  filter(STUSPS %in% c("WA", "OR", "CA", "ID", "NV", "AZ", "MT", "WY",
                       "UT", "CO", "NM", "ND", "SD", "NE", "KS", "OK", "TX",
                       "MN", "IA", "MO", "AR", "LA")) %>%
  st_geometry() %>%
  st_transform(crs_nad83)
rm(study_region_sf)
```


# Extreme wildfire risk: KBDI ≥ 600

## Descriptive statistics, aggregate

N wells in high wildfire risk areas (i.e., areas with KBDI ≥ 600):

- 2017: `r wells_kbdi %>% filter(kbdi_max_2017 >= 600) %>% nrow`
- 2050: `r wells_kbdi %>% filter(kbdi_max_2050 >= 600) %>% nrow`
- 2090: `r wells_kbdi %>% filter(kbdi_max_2090 >= 600) %>% nrow`


## Results, 2017 wildfire risk

**Plot. KBDI in study region**

```{r echo=FALSE}
ggplot() +
  geom_sf(data = kbdi_max_2017_sf, aes(color = kbdi_max_2017)) +
  scale_color_viridis(option = "B") +
  geom_sf(data = us_states_west, color = "white", fill = NA) +
  labs(color = "KBDI") +
  theme_minimal()
```


**Plot. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2017 >= 600) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  ggplot() +
  geom_bar(aes(state, n), stat = "identity") +
  theme_classic()
```

**Table. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2017 >= 600) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by county**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2017 >= 600) %>% 
  group_by(county_parish) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by operator**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2017 >= 600) %>% 
  group_by(operator) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```


## Results, 2050 wildfire risk

**Plot. KBDI in study region**

```{r echo=FALSE}
ggplot() +
  geom_sf(data = kbdi_max_2050_sf, aes(color = kbdi_max_2050)) +
  scale_color_viridis(option = "B") +
  geom_sf(data = us_states_west, color = "white", fill = NA) +
  labs(color = "KBDI") +
  theme_minimal()
```

**Plot. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2050 >= 600) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  ggplot() +
  geom_bar(aes(state, n), stat = "identity") +
  theme_classic()
```

**Table. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2050 >= 600) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by county**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2050 >= 600) %>% 
  group_by(county_parish) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by operator**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2050 >= 600) %>% 
  group_by(operator) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```


## Results, 2090 wildfire risk

**Plot. KBDI in study region**

```{r echo=FALSE}
ggplot() +
  geom_sf(data = kbdi_max_2090_sf, aes(color = kbdi_max_2090)) +
  scale_color_viridis(option = "B") +
  geom_sf(data = us_states_west, color = "white", fill = NA) +
  labs(color = "KBDI") +
  theme_minimal()
```

**Plot. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2090 >= 600) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  ggplot() +
  geom_bar(aes(state, n), stat = "identity") +
  theme_classic()
```

**Table. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2090 >= 600) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by county**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2090 >= 600) %>% 
  group_by(county_parish) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by operator**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2090 >= 600) %>% 
  group_by(operator) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

# High wildfire risk: KBDI ≥ 450

## Descriptive statistics, aggregate

N wells in high wildfire risk areas (i.e., areas with KBDI ≥ 450):

- 2017: `r wells_kbdi %>% filter(kbdi_max_2017 >= 450) %>% nrow`
- 2050: `r wells_kbdi %>% filter(kbdi_max_2050 >= 450) %>% nrow`
- 2090: `r wells_kbdi %>% filter(kbdi_max_2090 >= 450) %>% nrow`


## Results, 2017 wildfire risk

**Plot. KBDI in study region**

```{r echo=FALSE}
ggplot() +
  geom_sf(data = kbdi_max_2017_sf, aes(color = kbdi_max_2017)) +
  scale_color_viridis(option = "B") +
  geom_sf(data = us_states_west, color = "white", fill = NA) +
  labs(color = "KBDI") +
  theme_minimal()
```

**Plot. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2017 >= 450) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  ggplot() +
  geom_bar(aes(state, n), stat = "identity") +
  theme_classic()
```

**Table. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2017 >= 450) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by county**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2017 >= 450) %>% 
  group_by(county_parish) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by operator**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2017 >= 450) %>% 
  group_by(operator) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

## Results, 2050 wildfire risk

**Plot. KBDI in study region**

```{r echo=FALSE}
ggplot() +
  geom_sf(data = kbdi_max_2050_sf, aes(color = kbdi_max_2050)) +
  scale_color_viridis(option = "B") +
  geom_sf(data = us_states_west, color = "white", fill = NA) +
  labs(color = "KBDI") +
  theme_minimal()
```

**Plot. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2050 >= 450) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  ggplot() +
  geom_bar(aes(state, n), stat = "identity") +
  theme_classic()
```

**Table. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2050 >= 450) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by county**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2050 >= 450) %>% 
  group_by(county_parish) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by operator**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2050 >= 450) %>% 
  group_by(operator) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```


## Results, 2090 wildfire risk

**Plot. KBDI in study region**

```{r echo=FALSE}
ggplot() +
  geom_sf(data = kbdi_max_2090_sf, aes(color = kbdi_max_2090)) +
  scale_color_viridis(option = "B") +
  geom_sf(data = us_states_west, color = "white", fill = NA) +
  labs(color = "KBDI") +
  theme_minimal()
```
**Plot. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2090 >= 450) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  ggplot() +
  geom_bar(aes(state, n), stat = "identity") +
  theme_classic()
```

**Table. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2090 >= 450) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by county**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2090 >= 450) %>% 
  group_by(county_parish) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by operator**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2090 >= 450) %>% 
  group_by(operator) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```


# High wildfire risk: KBDI ≥ 400

## Descriptive statistics, aggregate

N wells in high wildfire risk areas (i.e., areas with KBDI ≥ 400):

- 2017: `r wells_kbdi %>% filter(kbdi_max_2017 >= 400) %>% nrow`
- 2050: `r wells_kbdi %>% filter(kbdi_max_2050 >= 400) %>% nrow`
- 2090: `r wells_kbdi %>% filter(kbdi_max_2090 >= 400) %>% nrow`


## Results, 2017 wildfire risk

**Plot. KBDI in study region**

```{r echo=FALSE}
ggplot() +
  geom_sf(data = kbdi_max_2017_sf, aes(color = kbdi_max_2017)) +
  scale_color_viridis(option = "B") +
  geom_sf(data = us_states_west, color = "white", fill = NA) +
  labs(color = "KBDI") +
  theme_minimal()
```

**Plot. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2017 >= 400) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  ggplot() +
  geom_bar(aes(state, n), stat = "identity") +
  theme_classic()
```

**Table. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2017 >= 400) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by county**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2017 >= 400) %>% 
  group_by(county_parish) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by operator**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2017 >= 400) %>% 
  group_by(operator) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

## Results, 2050 wildfire risk

**Plot. KBDI in study region**

```{r echo=FALSE}
ggplot() +
  geom_sf(data = kbdi_max_2050_sf, aes(color = kbdi_max_2050)) +
  scale_color_viridis(option = "B") +
  geom_sf(data = us_states_west, color = "white", fill = NA) +
  labs(color = "KBDI") +
  theme_minimal()
```

**Plot. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2050 >= 400) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  ggplot() +
  geom_bar(aes(state, n), stat = "identity") +
  theme_classic()
```

**Table. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2050 >= 400) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by county**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2050 >= 400) %>% 
  group_by(county_parish) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by operator**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2050 >= 400) %>% 
  group_by(operator) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```


## Results, 2090 wildfire risk

**Plot. KBDI in study region**

```{r echo=FALSE}
ggplot() +
  geom_sf(data = kbdi_max_2090_sf, aes(color = kbdi_max_2090)) +
  scale_color_viridis(option = "B") +
  geom_sf(data = us_states_west, color = "white", fill = NA) +
  labs(color = "KBDI") +
  theme_minimal()
```
**Plot. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2090 >= 400) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  ggplot() +
  geom_bar(aes(state, n), stat = "identity") +
  theme_classic()
```

**Table. Wells in high wildfire risk areas by state**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2090 >= 400) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by county**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2090 >= 400) %>% 
  group_by(county_parish) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

**Table. Wells in high wildfire risk areas by operator**

```{r echo=FALSE}
wells_kbdi %>% 
  filter(kbdi_max_2090 >= 400) %>% 
  group_by(operator) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

