---
title: "Results - Wells by state and wildfire-year (primary analysis)"
output:
  html_notebook:
    theme: flatly
    highlight: haddock 
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
source("../../code/0-setup/01-setup.R")
```

```{r data_input, include=FALSE}
wells_wildfire_state_year <- 
  readRDS("../../data/processed/wells_wildfire_state_year.rds")
```

# Description

These analyses are for the assessments of the number of wells in all wildfire burn areas by state and year (1984-2020). These are oil and gas wells with any operation date before the year of the wildfire and wells with missing operation dates, which we presume to have been drilled before the study period in an era of poor record-keeping. In a sensitivity analysis, we omit wells without data on operation dates.

# Aggregated results for all states

## Wells in wildfire burn areas

### Total wells burned by decade
- 1984-1989: **`r wells_wildfire_state_year %>% filter(year %in% c(1984:1989)) %>% select(n_wells) %>% sum()`**
- 1990-1999: **`r wells_wildfire_state_year %>% filter(year %in% c(1990:1999)) %>% select(n_wells) %>% sum()`**
- 2000-2009: **`r wells_wildfire_state_year %>% filter(year %in% c(2000:2009)) %>% select(n_wells) %>% sum()`**
- 2010-2019: **`r wells_wildfire_state_year %>% filter(year %in% c(2010:2019)) %>% select(n_wells) %>% sum()`**

### Average annual wells burned by decade
- 1984-1989: **`r wells_wildfire_state_year %>% filter(year %in% c(1984:1989)) %>% select(n_wells) %>% sum() / 6`**
- 1990-1999: **`r wells_wildfire_state_year %>% filter(year %in% c(1990:1999)) %>% select(n_wells) %>% sum() / 10`**
- 2000-2009: **`r wells_wildfire_state_year %>% filter(year %in% c(2000:2009)) %>% select(n_wells) %>% sum() / 10`**
- 2010-2019: **`r wells_wildfire_state_year %>% filter(year %in% c(2010:2019)) %>% select(n_wells) %>% sum() / 10`**

### Table. Wells in wildfire burn areas by year

```{r}
wells_wildfire_state_year %>% 
  filter(year %in% c(1984:2020)) %>% 
  group_by(year) %>% 
  summarize(n = sum(n_wells),
            n_wells_buffer_1km = sum(n_wells_buffer_1km))
```

### Plot. Wells in wildfire burn areas by year

```{r echo=FALSE}
wells_wildfire_state_year %>% 
  group_by(year) %>% 
  summarize(n = sum(n_wells)) %>% 
  ggplot() +
  geom_bar(aes(year, n), stat = "identity") +
  labs(x = "Year", y = "Wells (n)") +
  theme_classic()
```

### Plot. Wells in wildfire burn areas by year with best fit line

```{r echo=FALSE}
wells_wildfire_state_year %>% 
  group_by(year) %>% 
  summarize(n = sum(n_wells)) %>% 
  ggplot(aes(year, n)) +
  geom_point() +
  geom_smooth(method = "loess", color = "green") +
  geom_smooth(method = "lm", color = "blue") +
  labs(x = "Year", y = "Wells (n)") +
  theme_classic()
```

### Model. Wells in burn area ~ year * state

```{r}
model_fit <- glm(n_wells ~ year + factor(state),
                 data = wells_wildfire_state_year)
summary(model_fit)
```

In models adjusted for state, the number of wells in wildfire burn areas increased throughout the study period, 1984-2020 (p = 0.001).


## Wells within 1 km of wildfire burn areas

### Plot. Wells within 1km of wildfire burn areas by year

```{r echo=FALSE}
wells_wildfire_state_year %>% 
  group_by(year) %>% 
  summarize(n = sum(n_wells_buffer_1km)) %>% 
  ggplot() +
  geom_bar(aes(year, n), stat = "identity") +
  labs(x = "Year", y = "Wells (n)") +
  theme_classic()
```

### Plot. Wells within 1 km of wildfire burn areas by year with best fit line

```{r echo=FALSE}
wells_wildfire_state_year %>% 
  group_by(year) %>% 
  summarize(n = sum(n_wells_buffer_1km)) %>% 
  ggplot(aes(year, n)) +
  geom_point() +
  geom_smooth(method = "loess", color = "green") +
  geom_smooth(method = "lm", color = "blue") +
  labs(x = "Year", y = "Wells (n)") +
  theme_classic()
```

### Model. Wells within 1km of burn area ~ year * state

```{r}
model_fit <- glm(n_wells_buffer_1km ~ year + factor(state),
                 data = wells_wildfire_state_year)
summary(model_fit)
```

In models adjusted for state, the number of wells within 1km of wildfire burn areas (including wells within and just outside of the burn areas) increased throughout the study period, 1984-2020 (p < 0.001).

# Stratified by state

## Wells in wildfire burn areas

### Table. Wells in wildfire burn areas by state

```{r echo=FALSE}
wells_wildfire_state_year %>% 
  group_by(state) %>% 
  summarize(n_wells_by_state = sum(n_wells),
            n_wells_buffer_1km_by_state = sum(n_wells_buffer_1km)) %>% 
  arrange(desc(n_wells_by_state))
```

### Plot. Wells in wildfire burn areas by state

```{r echo=FALSE}
wells_wildfire_state_year %>% 
  group_by(state) %>% 
  summarize(n = sum(n_wells)) %>% 
  ggplot() +
  geom_bar(aes(state, n), stat = "identity") +
  labs(x = "State", y = "Wells (n)") +
  theme_classic()
```

### Table. Wells in wildfire burn areas by state and year

```{r}
wells_wildfire_state_year %>% 
  filter(year %in% c(1984:2020)) %>% 
  group_by(state, year) %>% 
  summarize(n = sum(n_wells),
            n_wells_buffer_1km = sum(n_wells_buffer_1km))
```

### Plot. Wells in wildfire burn areas by state and year

```{r echo=FALSE}
wells_wildfire_state_year %>% 
  group_by(state, year) %>% 
  summarize(n = sum(n_wells)) %>% 
  ggplot() +
  geom_bar(aes(year, n), stat = "identity") +
  labs(x = "Year", y = "Wells (n)") +
  facet_wrap(. ~ state) +
  theme_classic()
```

### Plot. Wells in wildfire burn areas by year with best fit line

```{r echo=FALSE}
wells_wildfire_state_year %>% 
  group_by(state, year) %>% 
  summarize(n = sum(n_wells)) %>% 
  ggplot(aes(year, n)) +
  geom_point() +
  geom_smooth(method = "loess", color = "green") +
  geom_smooth(method = "lm", color = "blue") +
  labs(x = "Year", y = "Wells (n)") +
  facet_wrap(. ~ state) +
  theme_classic()
```

### Model. Wells in burn area ~ year * state (CA)

```{r}
model_fit <- glm(n_wells ~ year,
                 data = subset(wells_wildfire_state_year, state == "CA"))
summary(model_fit)
```

No time trend in California.

### Model. Wells in burn area ~ year * state (OK)

```{r}
model_fit <- glm(n_wells ~ year,
                 data = subset(wells_wildfire_state_year, state == "OK"))
summary(model_fit)
```

Increasing number of wells in burn areas in Oklahoma.
